Modelling the gas kinematics of an atypical Lyman alpha emitting compact dwarf galaxy

### 2017-06-29 paper/paper.tex
comment on enclosed mass

Star-forming Compact Dwarf Galaxies (CDGs) 
resemble the expected pristine conditions of the first galaxies in the
Universe and are the best systems to test models on primordial galaxy
formation and evolution. 
Here we report on one of such CDGs, \tol, which presents
a broad, single peaked, highly symmetric \lya\ emission line that had
evaded theoretical interpretation so far.  
In this paper we explain these features by two different physically
motivated kinematic models: 
an interstellar medium composed by outflowing clumps with 
random motions and an homogeneous gaseous sphere undergoing solid body
rotation.
It is the first time that an observed \lya\ spectrum can be explained
assuming either of these kinematic conditions.
The multiphase model requires a clump velocity dispersion of
at least \sigmaclump\ with outflows of \inftyclump\, while the
bulk rotation velocity is constrained to be $348^{+75}_{-48}$ \kms.
We argue that the multiphase model has more chances to be correct.
The clump velocity dispersion implies a dynamical mass of at least $2\times
10^{9}$ M$_{\odot}$, ten times its baryonic mass, which can be
explained if \tol\ is hosted by a dark matter halo of at least
$10^{11}$ M$_{\odot}$.
If future kinematic maps of \tol\ confirm the velocities suggested by the
multiphase model, it would provide additional support to expect this
condition to be present  in primordial \lya\ emitting galaxies.  
\end{abstract}

@@

\cite{2011ApJ...726..108T} showed that  that such dynamical
mass can be explained if a dE is hosted by a dark matter halo of at
least $10^{11}$ M$_{\odot}$ in mass. 
This estimate is based on computing the integrated mass profile of a
spherical dark matter halo with a Navarro-Frenk-White profile with its
concentration following the median mass-concentration
relationship found in the Bolshoi simulation
\citep{2012MNRAS.423.3018P}.

These results would imply that \tol\ is part of the dE
class. 
However, there are two puzzles that are opened up by this hypothesis. 
First, dE galaxies are found either in clusters or groups of
galaxies; while \tol\ does not belong either to a group or a cluster
\citep{2001AJ....121..169F}. 
The second issue is that, unlike a dE, \tol\ hosts a bright and compact
startburst; imaging and spectroscopy studies suggest that the
startburst started less than 4 Myr ago, while the low surface
brightness regins are consistent with an age below one Gyr
\citep{2001AJ....121..169F}.

@@

>>>>

Star-forming Compact Dwarf Galaxies (CDGs) 
resemble the expected pristine conditions of the first galaxies in the
Universe and are the best systems to test models on primordial galaxy
formation and evolution. 
Here we report on one of such CDGs, \tol, which presents
a broad, single peaked, highly symmetric \lya\ emission line that had
evaded theoretical interpretation so far.  
In this paper we explain these features by two different physically
motivated kinematic models: 
an interstellar medium composed by outflowing clumps with 
random motions and an homogeneous gaseous sphere undergoing solid body
rotation.
It is the first time that an observed \lya\ spectrum can be explained
assuming either of these kinematic conditions.
The multiphase model requires a clump velocity dispersion of
at least \sigmaclump\ with outflows of \inftyclump\, while the
bulk rotation velocity is constrained to be $348^{+75}_{-48}$ \kms.
We argue that the multiphase model has more chances to be correct.
The clump velocity dispersion implies a dynamical mass of at least $2\times
10^{9}$ M$_{\odot}$, ten times its baryonic mass, which can be
explained if \tol\ is hosted by a dark matter halo of at least
$6\times 10^{11}$ M$_{\odot}$.
If future kinematic maps of \tol\ confirm the velocities suggested by the
multiphase model, it would provide additional support to expect this
condition to be present  in primordial \lya\ emitting galaxies.  
\end{abstract}

@@

Figure \ref{fig:mass} illustrates  how such dynamical
mass can be explained if \tol\ is hosted by a dark matter halo of at
least $6\times 10^{11}$ M$_{\odot}$ in virial mass. 
This estimate is based on computing the integrated mass profile of a
spherical dark matter halo with a Navarro-Frenk-White profile with its
concentration following the median mass-concentration
relationship found in the Bolshoi simulation
\citep{2012MNRAS.423.3018P}.

These results would imply that \tol\ is part of the dE
class. 
However, this hypothesis opens up two puzzles.
First, dE galaxies are found either in clusters or groups of
galaxies; while \tol\ does not belong either to a group or a cluster
\citep{2001AJ....121..169F}. 
The second issue is that, unlike a dE, \tol\ hosts a bright and compact
startburst; imaging and spectroscopy studies suggest that the
startburst started less than 4 Myr ago, while the low surface
brightness regins are consistent with an age below one Gyr
\citep{2001AJ....121..169F}.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.46\textwidth]{enclosed_mass.pdf}
\caption{Enclosed mass in a spherical system as a function of radius.
  Lines correspond to the expectation for dark matter halos and the point
  to the dynamical mass lower bounds for \tol.
  The dynamical mass estimates for \tol\ are consistent with the
  galaxy being hosted by a dark matter halo of at least
  $6\times10^{11}$ M$_{\odot}$ in mass.
    \label{fig:mass}} 
\end{center}
\end{figure}
